An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms

### 2022-11-17 Manuscript/manuscript_draft_ALI.tex
10/17/22 commit

\documentclass[12pt,a4paper]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{multirow}
\usepackage{multicol}
\setlength{\columnsep}{1.5cm}
\usepackage{setspace}
\usepackage{url}
\usepackage{array}
\usepackage{xfrac}
\usepackage{mwe}
\usepackage{tikz}
\usepackage{lmodern}
\usepackage{flowchart}
\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{color,soul}
%\usepackage[numbers,sort&compress]{natbib}
\usepackage{hyperref}
%\usepackage{cite}
\usepackage{biblatex}
\addbibresource{References.bib}
\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}

\title{Thank You, Next}
\author{Ali M. Campbell\\
	Tom E.X. Miller}
\begin{document}
	\maketitle

@@

Mutualisms are among the most widespread species interactions with diverse and dynamic consequences . Mutualisms are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity. Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies. It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms, portfolio effect, complementarity, and sampling effect. In this paper, I will focus on defensive ant-plant mutualisms. These involve plants which provide extra-floral nectar and/or “housing” to ants which in turn defend them from herbivores. While these interactions have been well studied in the literature, few have considered how diversity within partner guilds affect the overall benefits of mutualism for the plant partner. 
I use the plant, Cylindropuntia imbracata (tree cholla), and ant, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), and more, multispecies mutualism in which the cacti provide extrafloral nectar in exchange for defense from various herbivores and seed predators. I used 18 years of data collected from demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, ant partner, and herbivory for 8 30x30 m plots, in New Mexico. With this data I parameterize a series of Bayesian generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. I found that different ant partners did have different impacts on the vital rates of the tree cholla. Specifically, Crem. tended plants had advantages in both growth and survival when small, and Liom. tended plants had advantages in colonizing partners as well as floral viability. With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest fitness, and which of the above mechanisms may explain the effects of partner diversity in this system. I found that the real-life scenario (all possible ants are present) lead to the highest fitness for the tree cholla, indicating that partner diversity is beneficial in this system. It also shows that complementarity is at play in this system, meaning different partners offer different benefits leading to synergistic benefits for the tree cholla.

\section*{Context/Introduction}
BASELINE OF MUTUALISMS
Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,Stachowicz2005,BoucherDouglasH.1985} with diverse and dynamic consequences. 
\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea? Do I need to include more basic mutualism background? I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}.

Historically, scientists focused on pairwise mutualisms\cite{Bascompote2019}, however, in recent years multi-species mutualisms have been more heavily studied. 
It has become apparent that partner diversity in mutualisms often leads to non-additive fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}. 
This means the benefits can be greater than, less than, or equal to the sum of the benefits observed in each of the pairwise interactions with the focal mutualist that makes up the multi-species mutualism. 
This is true for many complicated reasons, including interactions between the various partners, niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability\cite{stanton2013, Afkhami2014,Boucher1982}.
Often partners have similar resource needs, particularly if they are in the same guild, which can lead to resource competition and fighting, potentially reducing benefits, or to increased quality and frequency of benefits\cite{Bascompote2019,Stanton2003}.
For this reason, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020} which are more common.
\ali{***** Biodiversity ecosystem function? }{After this I feel like I go straight into the mechanisms which explain the potential positive effects of diversity, but I don't specify that I am talking about those and I also don't talk about any explanations of negative effects? Maybe if I include the connection between BEF theory and mutualisms here it will make a nice bridge to the benefits of diversity?} 
Observational studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.

The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \{Stanton2003}.
Partner quality variations can remain consistent across years and life stages of the focal mutualist or they can shift year to year. 
When partner quality remains consistent across time, meaning one partner is always better or worse than another, the effects of partner diversity on the focal mutualist can be explained by a mechanism called sampling effect.
This is the concept that if partners vary in quality, then a more diverse sample of the partner community may be more likely to include the ``best" one\cite{Batstone2018}.
When a focal mutualist in a multi-species mutualism \ali{experiences}{not sure if this is the right word} sampling effect the fitness of that focal mutualist should be equal to the fitness of the focal mutualist if they interacted only with the ``best" partner. 
In some cases the quality of partners is not constant, rather it varies year to year due to environmental limitations, such as resource shortages and niche shifts.
The effects of partner diversity in these systems can be explained by the portfolio effect, the idea that each partner is the ``best" under a different set of conditions, which vary temporally\cite{Winfree2020,Batstone2018}. 
This mechanism works almost like a financial portfolio, when one partner has a bad year another partner has a good year, meaning the overall fitness benefits of the interactions remain relatively constant across temporal variation\cite{Batstone2018}.
Finally, partners can vary functionally, meaning each partner offers a different type of reward to the focal mutualist. 
The effects of partner functional diversity can be explained by the mechanism complementarity, the idea that each partner may account for some part of the benefits, but each has different functional pathways, which together can aid in broader benefits\cite{Winfree2020,Batstone2018}.
This variation can be both within and outside of a guild, meaning the variation can be very large, such as the difference between pollination and defense among other functions\cite{Bronstein2006}, or relatively small, such as similar rewards across different life stages\{Stanton2003}. 
Complementarity in a system is evidenced by synergistic benefits, meaning the fitness of the focal mutualist is higher than the fitness if the focal mutualist interacted with any one ``best" partner\{Batstone2018}. 
\ali{These three mechanism could explain the positive partner diversity effects on a focal mutualist in a multi-species mutualism, if these benefits exist.}{I am struggling on the transition between this paragraph and the next.}

Multi-species mutualisms are a huge group of interactions which vary considerably. 
While there are many types of mutualistic interactions, including pollination, defense, dispersal, housing, and many more\cite{Bronstein2006} in this paper we will focus on defensive ant-plant mutualisms.
These involve plants which provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
While these interactions have been well studied\cite{} in the literature, few have considered how diversity within ant defender guilds affect the overall benefits of mutualism for the plant partner.

This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla) and its ant mutualists. 
Tree cholla have a range which extends across the Southwest United States into Northern Mexico, with our study area set in New Mexico, US at the Sevilleta LTER.

These cacti flower in the spring throughout the summer (May to August), their reproductive season, outside of which they are not tended by ants. 
During the reproductive season these plants experience increased visitation by seed predators and herbivores, which influence their reproductive outputs, survival, growth, etc.
There are a number of observed ant partners which interact with the cacti in this part of their range, including \textit{Crematogaster opuntiae} (\textit{Crem.}), \textit{Liometopum apiculatum} (\textit{Liom.}), \textit{Forelius pruinosus}, an unidentified \textit{Camponotus} species, an unidentified \textit{Phenogaster} species, among others. 
The ants which most commonly interact with the cacti are \textit{Liom.} and \textit{Crem.}, while others are seen with low enough frequency that for the purpose of this paper we refer to them as ``other" ants. 
These ant partners are all in the same guild, meaning they all provide denfense to the cacti from insect herbivores and seed predators in return for extrafloral nectar (EFN).
The cacti in this mutualism are each tended by one species at a time, for their entire reproductive season with little turnover until the winter between reproductive seasons. 
Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.

\subsection{Questions}
In this paper we will answer several questions about the demographic effects of partner diversity in multi-species mutualisms:
\begin{enumerate}
  \item{What are the contrasting demographic effects of multiple partners?}
  \item{What combination of ant partners is associated with the highest fitness for the tree cholla?}
  item{What mechanisms explain the effects of multiple ant partners on the tree cholla?}
\end{enumerate}

To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti. 
We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
With the IPM we moved beyond what has been done in previous studies\cite{Palmer2010} to identify which mechanisms explain the effects of partner diversity on the tree cholla.

We hypothesized that the cacti would experience the highest fitness when interacting with all partners observed in the field, meaning the current mutualism results in the highest fitness for the cacti.
We also hypothesized that complementarity would explain these positive effects of partner diversity in the multi-species mutualism.

\section*{Methods}
\subsection*{Study System}
\subsection*{Natural History}
To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
These arborescent cacti produce cylindrical segments with large spines. 
In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
While most plants produce new segments every season, only those which are mature enough produce flower buds. 
Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.

This EFN is harvested by various ant species in return for defense. 
At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) } and \textit{Camponotus spp. (Camp.)}.

\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year \ref{Fig:Ant_freq}. 
It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.

These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next. 
Across the entire lifespan cacti may interact with multiple species of partners. 
At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\{Miller2014}. 
Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed. 
\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.

There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}. 
More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
These include an unidentified weevil, a cactus bug, and a seed predator. 
The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
The cactus bug, \textit{Narnia pallidicornis, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary. 
These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller 2009}.

@@

The data collected on these cacti are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
Six of the 30 $\times$ 30 meter plots were created with naturally occurring plants, censused from 2009 to 2014. 
The final two plots were added to this census from 2011 onwards. 
The distribution of these eight plots are shown in figure \ref{fig:map.
Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
For each plant, we recorded the height (cm), maximum crown width, and perpendicular crown width, which are used to calculate plant volume (cm3) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter. 
We recorded plant survival from the last survey to the current survey as a binary data point, 0 (died) or 1 (survived). 
We recorded the total number of flower buds, including how many were aborted and how many were not, which we used to calculate the viability rates of the cacti. 
We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
We also recorded the species of herbivore or seed predator, the number present on the plant, and if there is any clear, new evidence of herbivory on the tree cholla segments.

In addition to this primary dataset, I used several other datasets throughout the analysis in this paper.  Look at how Tom has described these. Cite these other papers which this data comes from. 
The second dataset was *** germination rates.
The third dataset was *** pre-census survival.
The fourth dataset was *** JO_fruit_data_final_dropplant0.csv.

@@

\vspace{1cm}

\section*{Results}
\subsection*{Vital Rates}
\includegraphics{"Figures/grow_countour_lines_color.png"}
The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:grow}.
Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
Plants tended by \textit{Crem.} ants had higher growth rates than tree cholla tended by any other ant or vacant. 
The next highest growth rate was seen in plants tended by other ants, followed by \textit{Liom.} ants, then vacant plants. 
The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
\ali{ }{I’m really unsure about what else I should say about this. }

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{multirow}
\usepackage{multicol}
\setlength{\columnsep}{1.5cm}
\usepackage{setspace}
\usepackage{url}
\usepackage{array}
\usepackage{xfrac}
\usepackage{mwe}
\usepackage{tikz}
\usepackage{lmodern}
\usepackage{flowchart}
\usetikzlibrary{shapes,arrows,positioning,calc,fit}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{color,soul}
\usepackage{hyperref}
\usepackage{biblatex}
\addbibresource{References.bib}
\newcommand{\tom}[2]{{\color{red}{#1}}\footnote{\textit{\color{red}{#2}}}}
\newcommand{\ali}[2]{{\color{pink}{#1}}\footnote{\textit{\color{pink}{#2}}}}

\title{An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms}
\author{Ali M. Campbell\\
	Tom E.X. Miller}
\begin{document}
	\maketitle

@@

Mutualisms are among the most widespread species interactions with diverse and dynamic consequences. They are considered more context dependent than other species interactions, meaning there are many different factors which influence the impacts of a mutualistic interaction, including partner diversity. Partner diversity has become a central focus in the field of mutualisms in recent years shifting the focus to multispecies mutualisms where the focus was previously on pairwise studies. It has been shown that pairwise studies are poor predictors of the effects of multispecies mutualistic interactions. The diversity of partners in a multi-species mutualism causes varied demographic effects on the population of the focal mutualist which can be explained by several mechanisms, portfolio effect, complementarity, and sampling effect. In this paper, I will focus on defensive ant-plant mutualisms. These involve plants which provide extra-floral nectar and/or “housing” to ants which in turn defend them from herbivores. While these interactions have been well studied in the literature, few have considered how diversity within partner guilds affect the overall benefits of mutualism for the plant partner. 
I use the plant, Cylindropuntia imbracata (tree cholla), and ant, Crematogaster opuntiae (Crem.), Liometopum apiculatum (Liom.), and more, multispecies mutualism in which the cacti provide extrafloral nectar in exchange for defense from various herbivores and seed predators. I used 18 years of data collected from demographic censuses, which includes data such as size, survival, reproductive status, flowers produced, ant partner, and herbivory for 8 30x30 m plots, in New Mexico. With this data I parameterize a series of Bayesian generalized linear vital rate models to determine the impacts of different partners on the focal mutualists. I found that different ant partners did have different impacts on the vital rates of the tree cholla. Specifically, Crem. tended plants had advantages in both growth and survival when small, and Liom. tended plants had advantages in colonizing partners as well as floral viability. With these models I constructed an Integral Projection Model in which I could vary the presence of each partner, creating different “diversity scenarios”, to determine under which diversity scenario the focal mutualist experienced the highest fitness, and which of the above mechanisms may explain the effects of partner diversity in this system. I found that the real-life scenario (all possible ants are present) lead to the highest fitness for the tree cholla, indicating that partner diversity is beneficial in this system. It also shows that complementarity is at play in this system, meaning different partners offer different benefits leading to synergistic benefits for the tree cholla.

\section*{Context/Introduction}

Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007}.
\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea? Do I need to include more basic mutualism background? I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}. 
While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species.
%Understanding why these relationships deteriorate or persist is important as mutualism is a cornerstone in maintaining the biodiversity of many systems\cite{Isbell2011}.

Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
%Interactions between partners, partner niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability all impact the fitness benefits to the focal mutualist\cite{stanton2013, Afkhami2014,Boucher1982}.
For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.

Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure ***, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}. 
Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure ***, through the portfolio effect\cite{Batstone2018,Lazaro2022}.

When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005}. 
Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
High degrees of partner turnover, also known as oportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
For example, susceptibility to enemies can change across life stages\cite{Boege2005}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.

Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms. 
While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{} affect the overall benefits of mutualism for the plant partner.

This study focuses on the long lived \textit{Cylindriopuntia imbricata} (tree cholla), long-lived EFN-bearing plants that associate with multiple species of ant partners.
Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
The ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
The multiple ant partners do not co-occur on plants, but a single cactus may interact with multiple species over its lifetime.
Some of the ant partners are more effective defenders than others, with \textit{Liom.} tended reproducing plants experiencing the lowest level of herbivory as shown in Figure \ref{fig:herb}.
This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.

\begin{figure}[h]
	\includegraphics[width=\linewidth]{herb_ant_only_flow.png}
	\caption{asdasdf}
	\label{fig:herb}
\end{figure}

In this paper we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
\begin{enumerate}
	\item{What are the contrasting demographic effects of multiple partners?}
	\item{What is the directionality of turnover between multiple partners?}
	\item{What mechanisms explain the benefits of partner diversity in a multispecies mutualism?}
\end{enumerate}

To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti. 
We analyzed the impacts of partner diversity on each individual vital rate (survival, growth, floral viability, etc.) with Bayesian models, which we used to parameterize an Integral Projection Model (IPM).
In previous studies the effects of multiple partners have been identified and quantified\cite{Palmer2010}, but none have identified the mechanisms that explain the benefits of partner diversity.

\section*{Methods}
\subsection*{Study System}

To determine how interactions with multiple partners impacts the cacti over time, we have monitored a population of tree cholla annually since 2004. 
This study was conducted in the Los Pinos mountains, a small mountain chain located on the Sevilleta National Wildlife Refuge, a Long-term Ecological Research (LTER) site in central New Mexico (34*20’5.3’’N, 106*37’53.2’’W).
This is an area characterized by steep, rocky slopes, perennial vegetation like cacti and junipers. 
The tree cholla cacti are common in high Chihuahuan desert habitats, with their native range spanning the southwestern USA\cite{Benson1982}. 
These arborescent cacti produce cylindrical segments with large spines. 
In the growing season, May to August in New Mexico, the plants initiate new vegetative segments and flower buds at the ends of existing segments. 
While most plants produce new segments every season, only those which are mature enough produce flower buds. 
Tree cholla generally reach at least 9 years of age before beginning to reproduce *** I found this being cited as unpublished data in Ohm Miller 2014 ***. 
Like other EFN-bearing cacti, tree cholla cacti secrete nectar from specialized glands on young vegetative or reproductive structures\cite{Ness2006,Oliveira1999}.

This EFN is harvested by various ant species in return for defense. 
At the Sevilleta, the cacti are visited primarily by two species of ground-nesting ants from the formicoid clade, \textit{Crematogaster opuntiae (Crem.) } and \textit{Liometopum apiculatum (Liom.), } as well as other rarer species, including \textit{Forelius pruinosus (For.) }, a \textit{Phenogaster} ant, and \textit{Camponotus spp. (Camp.)}.

\begin{figure}[h]
	\includegraphics[width=\linewidth]{Timeseries.png}
	\caption{asdfasdfsadfasdf}
	\label{fig:timeseries}
\end{figure}

\textit{Liom.} ants were the most frequent visitors, present a  between 25\% - 60\% of cacti annually, \textit{Crem.} were present at between 0\% - 20\% of cacti, and \textit{For.} were present at between 0\% - 5\% of cacti in any given year, as shown in Figure \ref{fig:timeseries}. 
It is important to note that not all cacti are visited every season, they can remain vacant, in fact up to 80\% of cacti are vacant in any given year.

These ants rarely co-occur on a plant, probably due to interspecific competition\cite{Miller2007}, rather each cactus is visited by a single ant species for the duration of a season, and the visitor can change from one season to the next. 
Across the entire lifespan cacti may interact with multiple species of partners. 
At the beginning of the growing season, when EFN production begins, the ground-nesting ants will begin visiting tree cholla.
They will visit the cacti every day during the season from early morning *** to sunset *** abandoning the cacti at night for their nests. 
In late August, the tree cholla will stop producing EFN and the ants will stop visiting until the next growing season. 
The process by which ants are sorted out onto different cacti is somewhat of a mystery, but we know several facts which likely influence the process. 
Smaller cacti are less likely to be visited because they produce very little EFN, so larger cacti are generally more desirable\cite{Miller2014}. 
Similarly, many factors go into how many vegetative segments and floral buds are produced in a particular growing season, so the more productive cacti are also more desirable. 
Some of the ant species are considered more aggressive, specifically \textit{Liom.}\cite{Miller2007}, which may be advantageous in going after more desirable cacti. 
Some ants are also more likely to retain the same tree cholla partners year after year, meaning they reduce the partner turnover experienced by individual cacti. 
These factors mean that by some hierarchy at the beginning of the season the ants claim a subset of the existing cacti and visit them regularly throughout the season, and that from season to season the same ant species may visit the same cacti, but it is not guaranteed. 
\ali{Note.}{I am not sure how much of this I should include here, since some of the turnover info is results from this paper?}
These ants defend the tree cholla from various seed predators and herbivores in exchange for the EFN.

There are a variety of insect herbivores and seed predators which attack the cacti, focusing either on the vegetative segments and the reproductive segments throughout the entire range of the tree cholla\cite{Mann1969}. 
More locally, at the Sevilleta, there are several insects which occur commonly on these cacti.
These include an unidentified weevil, a cactus bug, and a seed predator. 
The unidentified weevil of the genus \textit{Gerstaekeria} which feed on vegetative and reproductive structures and implant their larvae within the plant for the winter. 
The cactus bug, \textit{Narnia pallidicornis}, (Hemiptera: Coreidae) feed on all cactus parts with a preference for the reproductive structures \cite{Miller2006}.
The seed predator, \textit{Cahela ponderosella}, (Lepidoptera: Pyralidae) attack developing fruits pre-dispersal and oviposit in open flowers mid-growing season where larvae burrow into the ripening ovary. 
These predators can have significant negative impacts on the fitness of individual cacti and depress population growth\cite{Miller2009}.

@@

The data collected on these cacti are from a long-term dataset spanning 2004 to 2022 taken from our eight 30 $\times$ 30 meter plots at the Sevilleta LTER. 
The data initially included 134 naturally occurring plants across 4 spatial blocks censused annually from 2004 to 2008.
Six of the 30 $\times$ 30 meter plots were created with naturally occurring plants, censused from 2009 to 2014. 
The final two plots were added to this census from 2011 onwards. 
The distribution of these eight plots are shown in figure \ref{fig:map}.
Annually, in May we surveyed these plots, taking many types of demographic and partner data. 
For each plant, we recorded the height (cm), maximum crown width, and perpendicular crown width, which are used to calculate plant volume (cm3) based on the volume of a cone with the mean of maximum crown width and perpendicular crown width as the diameter. 
We recorded plant survival from the last survey to the current survey as a binary data point, 0 (died) or 1 (survived). 
We recorded the total number of flower buds, including how many were aborted and how many were not, which we used to calculate the viability rates of the cacti. 
We recorded all ant species present, usually just one, and the number of ants we could count in 30 seconds. 
We also recorded the species of herbivore or seed predator, the number present on the plant, and if there is any clear, new evidence of herbivory on the tree cholla segments.

In addition to this primary dataset, I used several other datasets throughout the analysis in this paper.  Look at how Tom has described these. Cite these other papers which this data comes from. 
The second dataset was *** germination rates.
The third dataset was *** pre-census survival.
The fourth dataset was *** $JO_fruit_data_final_dropplant0.csv$.

\begin{figure}[h]
%	\includegraphics[width=\linewidth]{map.png}
	\caption{asdfasdf}
	\label{fig:map}
\end{figure}

@@

\section*{Results}
I’m honestly not sure what all should be included here. I feel like I need a lot more info about the specifics???

\subsection*{Vital Rates}
The results of the Bayesian growth model show the probability of a cactus being a specific size given the size in the previous year and the ant partner in Figure \ref{fig:growth}.
Our analyses showed that plants of a specific size were more likely to grow larger by the next year if tended by certain ant partners. 
Plants tended by \textit{Crem.} ants had higher growth rates than tree cholla tended by any other ant or vacant. 
The next highest growth rate was seen in plants tended by other ants, followed by \textit{Liom.} ants, then vacant plants. 
The standard deviation of the growth rates also vary with the size of the tree cholla, decreasing as the size of the plants increase, ranging from $-1.089$ to $1.499$.
\ali{ }{I’m really unsure about what else I should say about this. }

\begin{figure}[h]
	\includegraphics[width=.5\linewidth]{grow_contour_lines_color.png}
	\caption{asdfasdf}
	\label{fig:growth}
\end{figure}

Our analyses of the bayesian hierarchical survival model showed that large cacti have higher survival rates than small, regardless of the ant partner. 
It also showed that small cacti had significantly different survival rates depending on the ant partner. 
Small cacti experienced highest survival rates when tended by \textit{Crem.} (near $70\%$) as compared to the lowest survival rates when tended by \textit{Liom.} or other ants (below $60\%$), as shown in Figure \ref{fig:survival}.
As tree cholla grow, the probabilities of survival increase no matter the partner to nearly $100\%$ with plants tended by \textit{Crem.} and \textit{Liom.} reaching maximum survival first and plants that are vacant reaching last. 
\textit{Crem.} tended plants have survival rates ranging from $68.379\%$ to $99.998\%$, with the rates increasing with the size of the cacti.  
\textit{Liom.} tended plants have survival rates ranging from $35.997\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
Other tended plants have survival rates ranging from $15.078\%$ to $99.999\%$, with the rates increasing with the size of the cacti. 
Vacant plants have survival rates ranging from $22.031\%$ to $99.647\%$, with the rates increasing with the size of the cacti.
